SMT-Boosted Security Types for Low-Level MPC

### 2025-04-24 esop25/slides.tex
checkpoint

  \begin{block}{Reconstructive Sharing}
    $$
    \begin{array}{lll}
      \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
      \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
      \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
      \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
      \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
      \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
    \end{array}
    $$
  \end{block}
  \begin{itemize}
  \item $\elab{\mesg{s1}}{2}, \elab{\mesg{s1}}{3}, \elab{\locflip}{1}$ are the \emph{shares} of $\sx{1}{1}$,
    etc.
  \item \emph{algebraically} $\sx{1}{1} \eop \elab{\mesg{s1}}{2} + \elab{\mesg{s1}}{3} + \elab{\locflip}{1}$.
  \item \emph{probabilistically} any 2 out of three shares are in a uniform random distribution.
  \end{itemize}

@@

  To sum the secrets \emph{in secret}:

  \begin{block}{Additive Homomorphism: Sum-of-Shares}
    $$
    \begin{array}{lll}
      \mx{ss1}{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
      \mx{ss2}{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
      \mx{ss3}{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} 
    \end{array}
    $$
  \end{block}

  Additive homorphism:
  \begin{itemize}
  \item \emph{Algebraically}, $\mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{1} \eop \sx{1}{1} + \sx{2}{2} + \sx{3}{3}$ 
  \item \emph{Probabilistically}, protocol step is noninteractive and each of $\mx{ss1}{i}$ remain in a uniform distribution.
  \end{itemize}

@@

@@

Letting $\prog$ be the above protocol:
$$
\toeq{\prog} \wedge \ttt{out} \eop \mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{1} \models\ \ttt{out} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
$$

>>>>

  \begin{block}{Reconstructive Sharing: Protocol $\prog_{\mathit{sum3}}$}
    $$
    \begin{array}{lll}
      \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
      \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
      \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
      \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
      \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
      \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
    \end{array}
    $$
  \end{block}
  $\elab{\mesg{s1}}{2}, \elab{\mesg{s1}}{3}, \elab{\locflip}{1}$ are the \cemph{secret shares} of $\sx{1}{1}$,
    etc.
  \begin{itemize}
  \item \emph{algebraically} $\sx{1}{1} \eop \elab{\mesg{s1}}{2} + \elab{\mesg{s1}}{3} + \elab{\locflip}{1}$.
  \item \emph{probabilistically} any 2 out of three shares are in a uniform random distribution.
  \end{itemize}

@@

  \begin{block}{Additive Homomorphism: Sum-of-Shares}     
  To sum the input secrets \emph{in secret} extend $\prog_{\mathit{sum3}}$:
    $$
    \begin{array}{lll}
      \mx{ss1}{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
      \mx{ss2}{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
      \mx{ss3}{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} 
    \end{array}
    $$
  \end{block}

  Additive homorphism-- each $\mx{ssi}{i}$ for $i \in \setit{1,2,3}$ is a share:
  \begin{itemize}
  \item \emph{Algebraically}, $\mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{3} \eop \sx{1}{1} + \sx{2}{2} + \sx{3}{3}$ 
  \item \emph{Probabilistically}, protocol step is noninteractive and each $\mx{ssi}{i}$ remains in a uniform distribution.
  \end{itemize}

@@

\section{Using SMT}

\begin{frame}{Using Satisfiability Modulo Finite Fields (SMFF)}

  SMFF supports equational constraint definitions $\eqs$ over a finite field $\mathcal{F}_{p}$
  with a mature solver, in the style of SMT.

  \medskip
  Overture protocols $\prog$ have a straightforward interpretation as SMFF constraints $\toeq{\prog}$ , e.g.,
  the first two lines of $\prog_{\mathit{sum3}}$ translate as:
  $$
  %\toeq{
  %  \begin{array}{c}
  %  \begin{array}{lll}
  %    \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
  %    \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} 
  %  \end{array}\\
  %  = \\
    \begin{array}{lll}
      \elab{\mesg{s1}}{2} &\eop& \elab{\secret{1}}{1} \fminus \elab{\locflip}{1} \fminus \elab{\flip{x}}{1}\quad \wedge \\ 
      \elab{\mesg{s1}}{3} &\eop& \elab{\flip{x}}{1} 
    \end{array}
    %\end{array}
    %}
  $$

@@

Letting $\prog$ be the above protocol:
$$
\begin{array}{c}
  \toeq{\prog} \wedge \ttt{out} \eop \mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{3}\\ \models\\
  \ttt{out} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
\end{array}
$$
